---
title: "Bayesian modeling and prediction for movies"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(gridExtra)
library(statsr)
library(BAS)
```

### Load data

```{r load-data}
load("movies.Rdata")
```


* * *

## Part 1: Data

```{r}
head(movies)
```




    1 pt for correct reasoning for generabizability – Answer should discuss whether random sampling was used. Learners might discuss any reservations, those should be well justified.

    1 pt for correct reasoning for causality – Answer should discuss whether random assignment was used.


* * *

## Part 2: Data manipulation

```{r}
movies <- movies %>% mutate(
  feature_film = if_else(title_type == 'Feature Film', "yes", "no"),
  drama = if_else(genre == 'Drama', "yes", "no"),
  mpaa_rating_R = if_else(mpaa_rating == "R", "yes", "no"),
  oscar_season = if_else(thtr_rel_month %in% c(10,11,12), "yes", "no"),
  summer_season = if_else(thtr_rel_month %in% c(5,6,7,8), "yes", "no")
)
```


```{r}
movies <- movies %>% select(
    audience_score, feature_film, drama, runtime, mpaa_rating_R, thtr_rel_year, oscar_season, summer_season,
    imdb_rating, imdb_num_votes, critics_score, best_pic_nom, best_pic_win, best_actor_win,
    best_actress_win, best_dir_win, top200_box
)
head(movies)
```


## Part 3: Exploratory data analysis

```{r}
p1 <- ggplot(movies, aes(x = feature_film, y = audience_score)) + geom_boxplot()
p2 <- ggplot(movies, aes(x = drama, y = audience_score)) + geom_boxplot()
p3 <- ggplot(movies, aes(x = mpaa_rating_R, y = audience_score)) + geom_boxplot()
p4 <- ggplot(movies, aes(x = oscar_season, y = audience_score)) + geom_boxplot()
p5 <- ggplot(movies, aes(x = summer_season, y = audience_score)) + geom_boxplot()
p6 <- ggplot(movies, aes(x = runtime, y = audience_score)) + geom_point()
p7 <- ggplot(movies, aes(x = thtr_rel_year, y = audience_score)) + geom_point()
p8 <- ggplot(movies, aes(x = imdb_rating, y = audience_score)) + geom_point()
p9 <- ggplot(movies, aes(x = imdb_num_votes, y = audience_score)) + geom_point()
p10 <- ggplot(movies, aes(x = critics_score, y = audience_score)) + geom_point()
p11 <- ggplot(movies, aes(x = best_pic_nom, y = audience_score)) + geom_boxplot()
p12 <- ggplot(movies, aes(x = best_pic_win, y = audience_score)) + geom_boxplot()
p13 <- ggplot(movies, aes(x = best_actor_win, y = audience_score)) + geom_boxplot()
p14 <- ggplot(movies, aes(x = best_actress_win, y = audience_score)) + geom_boxplot()
p15 <- ggplot(movies, aes(x = best_dir_win, y = audience_score)) + geom_boxplot()
p16 <- ggplot(movies, aes(x = top200_box, y = audience_score)) + geom_boxplot()
grid.arrange(p1,p2,p3,p4)
grid.arrange(p5,p6,p7,p8)
grid.arrange(p9,p10,p11,p12)
grid.arrange(p13,p14,p15,p16)
```

```{r}
movies %>% group_by(feature_film) %>% summarise(mean = mean(audience_score), var = var(audience_score), cnt = n())
```

```{r}
movies %>% group_by(drama) %>% summarise(mean = mean(audience_score), var = var(audience_score), cnt = n())
```

```{r}
movies %>% group_by(mpaa_rating_R) %>% summarise(mean = mean(audience_score), var = var(audience_score), cnt = n())
```

```{r}
movies %>% group_by(oscar_season) %>% summarise(mean = mean(audience_score), var = var(audience_score), cnt = n())
```

```{r}
movies %>% group_by(summer_season) %>% summarise(mean = mean(audience_score), var = var(audience_score), cnt = n())
```

```{r}
movies_numeric <- movies %>% select(audience_score, runtime, thtr_rel_year, imdb_rating, imdb_num_votes, critics_score)
cor(movies_numeric)
```


```{r}
movies %>% group_by(best_pic_nom) %>% summarise(mean = mean(audience_score), var = var(audience_score), cnt = n())
```

```{r}
movies %>% group_by(best_pic_win) %>% summarise(mean = mean(audience_score), var = var(audience_score), cnt = n())
```

```{r}
movies %>% group_by(best_actor_win) %>% summarise(mean = mean(audience_score), var = var(audience_score), cnt = n())
```

```{r}
movies %>% group_by(best_actress_win) %>% summarise(mean = mean(audience_score), var = var(audience_score), cnt = n())
```

```{r}
movies %>% group_by(best_dir_win) %>% summarise(mean = mean(audience_score), var = var(audience_score), cnt = n())
```

```{r}
movies %>% group_by(top200_box) %>% summarise(mean = mean(audience_score), var = var(audience_score), cnt = n())
```



* * *

## Part 4: Modeling

imdb_rating: Rating on IMDB
critics_score: Critics score on Rotten Tomatoes
--
audience_score: Audience score on Rotten Tomatoes

model selection:
- best predictive
- highest post prob
- BIC

?

```{r}
fit <- bas.lm(
  data = movies,
  formula = audience_score ~ . ,
  prior = "ZS-null",
  #prior = "BIC",
  #prior = "AIC",
  modelprior = uniform(),
  #modelprior = beta.binomial(1,1),
  method = "BAS"
  #method = "MCMC"
  )   
```

```{r}
fit
```

```{r}
summary(fit)
```


```{r}
plot(fit, which = 1) 
```

```{r}
fit 
```

```{r}
fit 
```

```{r}
fit 
```

```{r}
fit 
```

* * *

## Part 5: Prediction

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button above. Make sure that your code is visible in the project you submit. 
Delete this note when before you submit your work.

* * *

## Part 6: Conclusion

