---
title: "Statistical Inference Course Project, Part 1"
author: "Sigrid Keydana"
output:
  html_document: default
---

```{r, echo = FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
```


# Central Limit Theorem

## Overview

First, we're going to demonstrate the Central Limit Theorem for samples of exponential data.

## Simulations

To demonstrate the CLT, we will sample 40 values from an exponential(0.2) distribution and do this a 1000 times.
```{r}
lambda <- 0.2
nsim <- 1000
n <- 40
```

We store the raw values ...

```{r}
c <- rexp(nsim * n, lambda)
m <- matrix(c, nsim, n)
dim(m)
```

... the means of each 40-element sample ...

```{r}
means <- apply(m, 1, mean)
head(means)
```

... as well as the mean of the sample means:

```{r}
sample_mean_of_means <- mean(means)
sample_mean_of_means
```


## Sample Mean versus Theoretical Mean

We can compare the means - and the mean of means - with the theoretical mean of the exponential(0.2) distribution, which is:

```{r}
dist_mean <- 1/lambda
dist_mean
```


Here we see a histogram of the 1000 means we obtained.
Also displayed is the theoretical mean (red line) and the mean of means (black line).

```{r}
ggplot(data_frame(m = means), aes(means)) + geom_histogram(bins = 15, color = 'cyan', fill = 'white') + 
  geom_vline(xintercept = dist_mean, color = 'red', size = 1) +  
  geom_vline(xintercept = sample_mean_of_means, color = 'black', size = 1) +
  ggtitle('Distribution of sample means from an exponential distribution, n=40')
```

As you see the theoretical mean and the mean of means are _very_ close.
Let's check the numbers:

```{r}
paste0('Mean of exponential distribution is: ', dist_mean, '. Mean of sample means is: ', sample_mean_of_means)
```


## Sample variance versus theoretical variance

Let's also have a look at the distribution of the sample variances.
We'll compare it to the theoretical variance of an exponential(0.2) distribution: 

```{r}
variances <- apply(m, 1, var)
head(variances)
```

In contrast to the means, the sample variances are _not_ normally distributed.

```{r}
dist_variance <- (1/lambda)^2
ggplot(data_frame(m = variances), aes(variances)) + geom_histogram(bins = 15, color = 'cyan', fill = 'white') + 
  geom_vline(xintercept = dist_variance, color = 'red', size = 1) +
  ggtitle('Distribution of sample variances from an exponential distribution, n=40')
```


## Distribution

Let's again have a look at the distribution of means, and compare it to the distribution of the raw data.
First, here is a histogram of the means again:

```{r}
ggplot(data_frame(m = means), aes(means)) + geom_histogram(bins = 15, color = 'cyan', fill = 'white') +
  ggtitle('Distribution of sample means from an exponential distribution, n=40')
```

Here, on the other hand, is a histogram of the raw data, with the empirical mean shown in black:

```{r}
whole_sample_mean <- mean(c)
ggplot(data_frame(c = c), aes(c)) + geom_histogram(bins = 15, color = 'cyan', fill = 'white') + 
  geom_vline(xintercept = whole_sample_mean, color = 'black', size = 1) +
  ggtitle('Distribution of exponentials, n=1000')
```

So we see that in contrast to the distribution of means, the exponential distribution of raw data is heavily skewed. We can also create quantile-quantile plots to check whether the means, and the raw data, respectively, are normally distributed. First, the means:

```{r}
qqnorm(means)
```

We see that theoretical quantiles and sample quantiles match very well. This is not at all the case for the raw data:

```{r}
qqnorm(c)
```


